---
title: "Tuango RFM HW"
author: "Rainy Chen"
date: "2/21/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(flextable)
library(gmodels)
library(Hmisc)
library(statar)
library(ggpubr)

TG <- read.csv("/Users/rarainy/Documents/Spring2022/MKT 500T/m5/Tuango_RFM.csv")
```
  
## Q1.
```{r}
CrossTable(TG$buyer, digits=4)
```
3.1% of customers responded after the push message.

## Q2.
```{r}
mean(TG$ordersize[TG$buyer==1])
```
Average spending is about 202 RMB. 

## Q3. 
```{r}
TG_Copy <- TG %>% 
  mutate(rec_quin=xtile(recency, 5),
         freq_quin=xtile(frequency, 5),
         mv_quin=xtile(monetary, 5))
head(TG_Copy %>% select(userid, buyer, recency, frequency, 
                    monetary, rec_quin, freq_quin, mv_quin))
```
## Q4.
```{r}
avg_resp_rate_rec <- TG_Copy %>% group_by(rec_quin) %>% 
  summarise(avg_resp_rate=mean(buyer), .groups="drop")
avg_resp_rate_rec
```
```{r}
bar_avg_resp_rate_rec <- 
  ggplot(data=avg_resp_rate_rec, 
         aes(x = rec_quin, y = avg_resp_rate)) + 
  labs(x="Recency Quintile", y="Average Response Rate") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_resp_rate_rec
```

## Q5. 
```{r}
avg_resp_rate_freq <- TG_Copy %>% 
  group_by(freq_quin) %>% 
  summarise(avg_resp_rate=mean(buyer), .groups="drop")
avg_resp_rate_freq
```
```{r}
bar_avg_resp_rate_freq <- 
  ggplot(data=avg_resp_rate_freq, 
         aes(x = freq_quin, y = avg_resp_rate)) + 
  labs(x="Frequency Quintile", y="Average Response Rate") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_resp_rate_freq
```

## Q6. 
```{r}
avg_resp_rate_mv <- TG_Copy %>% group_by(mv_quin) %>%
  summarise(avg_resp_rate=mean(buyer), .groups="drop")
avg_resp_rate_mv
# The bar chart
bar_avg_resp_rate_mv <- 
  ggplot(data=avg_resp_rate_mv, 
         aes(x = mv_quin, y = avg_resp_rate)) + 
  labs(x="Monetary Quintile", y="Average Response Rate") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_resp_rate_mv
```

## Q7. 
```{r}
#recency
avg_spend_rec <- TG_Copy %>% group_by(rec_quin) %>% 
  summarise(avg_spend=mean(ordersize[buyer == 1]), .groups="drop")
avg_spend_rec

#bar chart 
bar_avg_spend_rec <- 
  ggplot(data=avg_spend_rec, 
         aes(x = rec_quin, y = avg_spend)) + 
  labs(x="Rencency Quintile", y="Average Spending") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_spend_rec
```
```{r}
#frequency 
avg_spend_freq <- TG_Copy %>% 
  group_by(freq_quin) %>% 
  summarise(avg_spend=mean(ordersize[buyer == 1]), .groups="drop")
avg_spend_freq

#bar chart 
bar_avg_spend_freq <- 
  ggplot(data=avg_spend_freq, 
         aes(x = freq_quin, y = avg_spend)) + 
  labs(x="Frequency Quintile", y="Average Spending") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_spend_freq
```
```{r}
#monetary 
avg_spend_mv <- TG_Copy %>% group_by(mv_quin) %>%
  summarise(avg_resp_rate=mean(ordersize[buyer == 1]), .groups="drop")
avg_spend_mv

# The bar chart
bar_avg_spend_mv <- 
  ggplot(data=avg_spend_mv, 
         aes(x = mv_quin, y = avg_resp_rate)) + 
  labs(x="Monetary Quintile", y="Average Response Rate") + 
  geom_bar(stat="identity", width=0.5)
bar_avg_spend_mv
```

## Q8.
Overall, the more recent, frequent purchases, and more average spending per deal a customer has, the more it responds. For customers who bought the deal, there is no clear trend on RFM. 

## Part 2 Q9.
```{r}
#marginal cost/average sales revenue=average breakeven response rate
breakeven=1.6/(0.5*mean(TG$ordersize[TG$buyer==1]))
breakeven
```
Average breakeven response rate is 1.6%.

## Q10.
```{r}
#(a) profit=[(total customer*average response rate)*average spending]*50% of sales revenue as fee - marginal cost*total customer
profit=264841*0.031*202.3565*0.5-1.6*264841
profit

#(b) return on marketing cost=profit/(marginal cost*total customer)
return=profit/(1.6*264841)
return
```

## Q11.
```{r}
#number of customers belonging to profitable cells
TG_Copy <- TG_Copy %>% 
  group_by(rfm1) %>% 
  mutate(avg_resp_rate_rfm=mean(buyer)) %>% ungroup()
TG_Copy <- TG_Copy %>%  mutate(pushto = avg_resp_rate_rfm > breakeven)

rfm1_group_size <- TG_Copy[TG_Copy$pushto,] %>% group_by(rfm1) %>% 
  summarise(group_size_rfm1 = n(), .groups="drop") 

#number of buyers belonging to profitable cells
rfm1_buyer_size <- TG_Copy[TG_Copy$pushto&TG_Copy$buyer==1,] %>% group_by(rfm1) %>% 
  summarise(buyer_size_rfm1 = n(), avg_revenue_rfm1 = mean(ordersize)) 
```

```{r}
# profit
rfm1_profit <- left_join(rfm1_group_size, rfm1_buyer_size)
# Calculated using average revenue in each rfm group
rfm1_profit$profit_rfm1 = (rfm1_profit$avg_revenue_rfm1 * rfm1_profit$buyer_size_rfm1 * 0.5 - 1.6 * rfm1_profit$group_size_rfm1)   * (264841 / 13939)
rfm1_profit$return_rfm1 = rfm1_profit$profit_rfm1 / (1.6 * rfm1_profit$group_size_rfm1)

rfm1_profit[,c("rfm1", "profit_rfm1", "return_rfm1")]
sum(rfm1_profit$profit_rfm1) - profit
```

## Q12. 
```{r}
TG_Copy <- TG_Copy %>% mutate(same_rfm = rfm1==rfm2) 
CrossTable(TG_Copy$same_rfm)
```
There are about 87.2% of same RFM index. Sequential RFM pros is that it provides a more even distribution RFM group; cons is that the index is hard to interpret since frequency and monetary rank dependent on recency rank. Independent RFM pros is that it is easy to interpret, but may provide empty bin because of uneven distribution. 






